Parsing Postfix log files

### 2008-04-01 doc/logparser-acronyms.tex
The remainder of Carl's simple corrections.  There are some more I need to talk to him about.

\newacronym{regex}{Regular Expression}{description={
    Regular expressions are a method of matching patterns in text that are
    explained in detail in~\cite{Wikipedia-regex, perlre} .
}}

>>>>

\newacronym{regex}{Regular Expression}{description={
    Regular expressions are a method of matching patterns in text that are
    explained in detail in~\cite{Wikipedia-regex, perlre}.
}}

### 2008-04-01 doc/logparser.tex
The remainder of Carl's simple corrections.  There are some more I need to talk to him about.

Warnings are issued for any log lines which are not parsed; no warnings are
issued for unparsed log lines while testing with the \numberOFlogFILES{}
test log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: short of examining each
of the 60,721,709 log lines and determining which \regex{} parsed it, there
is no way to be sure that every line was parsed by the correct \regex{},
making it impossible to quantify the false positive rate; however a random
sample of 6039 log lines was parsed and the results checked manually to
ensure that the correct \regex{} parsed each log line.  The sample was
generated by running the following command:

@@

At the time of writing the parser is being tested with \numberOFlogFILES{}
log files.  There are 5 warnings produced, but because the parser errs on
the side of producing more warnings rather than fewer those 5 warnings
represent 3 instances of 1 problem: 3 connections started before the first
log file, so their initial log entries are missing, leading to warnings
when their log lines are parsed.

@@

% XXX SHOULD I HAVE THIS \newpage?
\newpage
\section{Conclusion}

@@

The division of the parser into rules, actions and framework is unusual
because rules are separated so completely from the actions and framework;
although parsers are often divided (lex and yacc~\cite{lex-and-yacc} based
parsers being an obvious example), the parts are usually quite internally
interdependent, and will be combined into a complete parser by the
compilation process; in contrast \parsername{} keeps the rules and actions
separate until the parser runs.  The actions and framework are not as
completely separated, as the actions depend on services provided by the
framework; however the actions and framework are not tightly integrated: it
would be possible, with some work, to completely separate the two.

@@

The real difficulties arise once the parser is successfully dealing with
90\% of the log lines, as the irregularities and complications previously
explained only begin to become apparent once the vast majority of log lines
have been parsed successfully.  Adding new rules to deal with numerous,
infrequently occurring log lines is a simple task, albeit tiresome (though
alleviated by the tool described in \sectionref{adding new actions}),
whereas dealing with mails where information appears to be missing is much
more grueling.  Trawling through the logs, looking for something out of the
ordinary, possibly hundreds or even thousands of lines away from the last
mention of the queueid in question, is extremely time consuming and error
prone.  In some cases the task is not to spot the line which is unusual,
but to spot that a line normally present is missing, i.e.\ to realise that
one line amongst hundreds is absent.  In all cases the evidence must be
used to construct a hypothesis to explain the irregularities, and that
hypothesis must then be tested in the parser; if successful the parser must
be modified to deal with the irregularities, without adversely affecting
the existing functionality.  The complications described in
\sectionref{additional complications} were solved in the order they are
described in, and that order closely resembles the frequency with which
they occur; the most frequently occurring complications dominate the
warning messages produced, and so naturally they are the first
complications to be dealt with.

@@

        Sawmill is a general purpose commercial product which parses 687
        log file formats (correct as of 2007/12/04) and produces reports.
        Its data extraction facilities are quite limited, though it does
        extract three different sets of data for Postfix (one is beta as of
        2007/12/04), but they do not appear to be interlinked, nor does it
        save sufficient data for the purposes of this project.  No attempt
        is made to correlate log lines or deal with the difficulties
        documented in \sectionref{complications} \sectionref{additional
        complications}.\footnote{If any attempt is made there is no
        reference to it in the documentation available on the website.} The
        source code is available in an obfuscated form only (presumably for
        a fee), and the product is quite expensive, as it requires a
        license per report which is to be generated; in contrast
        \parsername{} is free and the code is freely available.  The web
        interface allows creation of dynamic reports based on any field,
        but due to the associated cost the author has not experimented with
        it.  Presumably as the company charges per report the user wishes
        to generate the data store (if there is one) is deliberately
        inaccessible and undocumented to prevent the user bypassing the
        program and generating their own reports.

@@

The following graphs show the impact that not caching compiled \regexes{}
has on parser performance: on typical log files the execution time when not
caching compiled \regexes{} is 500--600\% of the execution time when
caching; reversing the perspective shows that cached execution time is
merely 17--20\% of non-cached execution time.  Caching compiled \regexes{}
is probably the single most effective optimisation possible in the parser's
implementation.

>>>>

Warnings are issued for any log lines which are not parsed; no warnings are
issued for unparsed log lines while testing with the \numberOFlogFILES{}
test log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: short of examining each
of the 60,721,709 log lines and determining which \regex{} parsed it, there
is no way to be sure that every line was parsed by the correct \regex{},
making it impossible to quantify the false positive rate; however a random
sample of 6039 log lines was parsed and the results checked manually to
ensure that the correct \regex{} parsed each log line.\footnote{Each log
line was examined and the correct \regex{} identified from the
\numberOFrules{} rules in the database; the correct \regex{} was then
compared to the \regex{} which was used by the parser.}  The sample was
generated by running the following command:

@@

At the time of writing the parser is being tested with \numberOFlogFILES{}
log files.  There are 5 warnings produced, but because the parser errs on
the side of producing more warnings rather than fewer, those 5 warnings
represent 3 instances of 1 problem: 3 connections started before the first
log file, so their initial log entries are missing, leading to warnings
when their log lines are parsed.

@@

\section{Conclusion}

@@

The division of the parser into rules, actions and framework is unusual
because rules are separated so completely from the actions and framework;
although parsers are often divided (parsers based on the combination of
\texttt{lex} and \texttt{yacc}~\cite{lex-and-yacc} 
being an obvious example), the parts are usually quite internally
interdependent, and will be combined into a complete parser by the
compilation process; in contrast \parsername{} keeps the rules and actions
separate until the parser runs.  The actions and framework are not as
completely separated, as the actions depend on services provided by the
framework; however the actions and framework are not tightly integrated: it
would be possible, with some work, to completely separate the two.

@@

The real difficulties arise once the parser is successfully dealing with
90\% of the log lines, as the irregularities and complications previously
explained only begin to become apparent once the vast majority of log lines
have been parsed successfully.  Adding new rules to deal with numerous,
infrequently occurring log lines is a simple task, albeit tiresome (though
alleviated by the tool described in \sectionref{adding new actions}),
whereas dealing with mails where information appears to be missing is much
more grueling.  Trawling through the logs, looking for something out of the
ordinary, possibly hundreds or even thousands of lines away from the last
mention of the queueid in question, is extremely time consuming and error
prone.  In some cases the task is not to spot the line which is unusual,
but to spot that a line normally present is missing, i.e.\ to realise that
one line amongst hundreds is absent.  In all cases the evidence must be
used to construct a hypothesis to explain the irregularities, and that
hypothesis must then be tested in the parser; if successful, the parser
must be modified to deal with the irregularities, without adversely
affecting the existing functionality.  The complications described in
\sectionref{additional complications} were solved in the order they are
described in, and that order closely resembles the frequency with which
they occur; the most frequently occurring complications dominate the
warning messages produced, and so naturally they are the first
complications to be dealt with.

@@

        Sawmill is a general purpose commercial product which parses 687
        log file formats (correct as of 2007/12/04) and produces reports.
        Its data extraction facilities are quite limited, though it does
        extract three different sets of data for Postfix (one is beta as of
        2007/12/04), but they do not appear to be interlinked, nor does it
        save sufficient data for the purposes of this project.  No attempt
        is made to correlate log lines or deal with the difficulties
        documented in \sectionref{complications} \sectionref{additional
        complications}.\footnote{If any attempt is made there is no
        reference to it in the documentation available on the website.} The
        source code is available in an obfuscated form only (presumably for
        a fee), and the product is quite expensive, as it requires a
        license per report which is to be generated; in contrast
        \parsername{} is free and the code is freely available.  The web
        interface allows creation of dynamic reports based on any field,
        but due to the associated cost the author has not experimented with
        it.  Presumably, as the company charges per report the user wishes
        to generate, the data store (if there is one) is deliberately
        inaccessible and undocumented to prevent the user bypassing the
        program and generating their own reports.

@@

The following graphs show the impact that not caching compiled \regexes{}
has on parser performance: on typical log files the execution time when not
caching compiled \regexes{} is 500--600\% of the execution time when
caching; reversing the perspective shows that cached execution time is
merely 17--20\% of non-cached execution time.  Caching compiled \regexes{}
is probably the single most effective optimisation possible in the parser's
implementation; given that it only required sixteen extra lines of
reasonably simple code,\footnote{The sixteen lines of code breaks down as
follows: three lines to add caching, eight lines of error checking and
reporting, three lines to optionally disable caching for debugging and
performance measurement, and two lines of comments.} the investment in time
was minimal.
